Supplemental Materials

### 2021-07-17 supplement/supplement-main.Rmd
supplement updates

---
title: "Supplemental Materials"
subtitle: "Untitled vaccination intuitive theory paper"
author: "Derek Powell, Kara Weisman, and Ellen M. Markman"
date: "June 21, 2021 (DRAFT)"
output:
  pdf_document: 
    latex_engine: xelatex
    fig_caption: yes
figsintext : yes
always_allow_html: true
header-includes:
 \usepackage{float}
---

@@

Figure 1 shows an example of a simple Bayesian Network. The network itself is a _directed acyclic graph_ (DAG), a type of graphical model. Graphical models represent variables as nodes connected by edges. In a DAG, these edges are directed, flowing from parent to child, and are constrained so that they cannot form cycles or loops. In a Bayesian network, these directed edges encode conditional dependencies among variables, so that the conditional probability distribution for a given variable is defined only in terms of its direct parents. Together, a child node and its parents are called a _family_.

```{r dag_example, echo=F, fig.align='center', out.width = '40%', fig.cap="Example DAG illustrating some variables related to wildfires."}

@@

In this graph, the edge $FIRE \rightarrow SMOKE$ captures that fire causes smoke.^[The example of cause and effect is especially intuitive, but these edges need not be causal. Instead, they are better thought of as representing “influence relations”, which may also include relationships like logical implication and set membership (Williamson, 2001).] This relationship is represented as a probabilistic function, where the probability of smoke depends on the probability of fire. The probability of a node Y can be defined from its parents X in many different ways. For binary nodes, the most flexible specification is a conditional probability table (CPT) with 2^n^ entries specifying the probability of Y given the entire joint probability distribution over X, where n is the number of parents X. Given the exponential complexity of the CPT, simpler functions are often used to specify the probability of the child node, such as the noisy-or and noisy-and-not functions. A noisy-or function assumes each of the parents are independent generative forces, each with a probability of generating the effect. A noisy-and-not function assumes each of the parents are independent preventive forces, specifying the probability of the child node as the probability that it is generated _and not_ prevented by any one of its possible preventers. We construct our cognitive model parameterization using a combination of these two functions, using a distribution sometimes referred to as a “DeMorgan gate” (Maaskant & Druzdel, 2008). In our model, the probability of the child node Y can be expressed as:

@@

A cognitive model composed of these types of relations is capable of capturing a causal system of non-interactive generative and preventive causes. Although people’s intuitive theories are likely more flexible than this, there is ample evidence that people are highly capable of reasoning according to these types of causal relationships (see Holyoak & Cheng, 2011 for a review), and moreover, that their causal learning is strongly biased toward learning these kinds of relations (e.g., Novick & Cheng, 2004).

@@

To conduct an efficient search through the super-exponential space of DAGs, we tested several variants of two structure learning algorithms implemented in the “bnlearn” R package. The first is a purely score-based “hill-climbing” algorithm (cite) and the second was a hybrid algorithm that combined the hill-climbing algorithm with a prior “restriction” phase that pruned down the space of DAGs. This pruning was accomplished with the “min-max parents and children” algorithm (cite), which can be tuned with a hyperparameter alpha, which we tested at several levels. To speed computations, we restricted the maximum number of parents in the graph to five for all nodes, as the complexity of the scoring function is exponential in the number of parents. In all resulting networks no node was found to have more than four parents, suggesting this did not impact the final results.

@@

Figure X presents the results of these cross validation tests. As seen in the figure, all structure learning approaches produced models with similar out-of-sample performance, with more variability among folds and runs than between algorithms. Thus the cross-validation does not uniquely identify an optimal structure learning strategy.

@@

```{r cross_validation, echo=FALSE, fig.align='center', fig.height=3, fig.width=4, fig.cap="Cross-validation performance for diffrent structure learning algorithms on the training data from Study 1."}

@@

We used a bootstrapping procedure to estimate the confidence in each of the links in this model, conducting the structure learning procedure repeatedly for 200 bootstrapped resamples of the original training data. Figure X visualizes the frequency with which each arc appeared in the resulting model across these replicates, and figure x visualizes the model with the coefficients for each link encoded by color.

@@

We can examine the robustness of the model’s features in two ways. First, we can ask about the “skeleton” of the graphical model: whether the relationships between pairs of nodes are robust, irrespective of their directions. To test this, we applied our chosen structure learning algorithm against 1000 bootstrapped replicates of our data (following Friedman et al., 2013). As shown in figure X, we have a high degree of confidence in the “skeleton” of the graph: nodes which are connected in the final graph are reliably connected across bootstrapping runs.

@@

```{r boot_strength_plot, echo=F, message=F, fig.cap="Bootstrapped confidence estimates for Bayesian network edges. Edge color indicates frequency with which edges appeared when conducting search across X bootstrap replicates. Only edges appearing in at least 20\\% of graphs are shown."}
# 
# hyde_str <- df_to_hyde_string(boot_res %>% filter(strength>.2, direction>0) %>% select(from,to))
# h2 <- HydeNetwork(as.formula(hyde_str))
# # plot(h2)
# 
# z2 <- custom_plot.HydeNetwork(h2)
# boot_res %>%
#   filter(direction>0, strength>.2) %>% 
#   # filter(parameter!="phi",parameter!="w_leak") %>%
#   # rename(from = parameter) %>%
#   # mutate(from = gsub("w_","",from)) %>%
#   mutate(strength = strength*direction) %>% 
#   select(to, from, strength) %>%
#   merge(z2$nodes_df %>% rename(from = label) %>% select(id, from), by="from") %>%
#   mutate(from = id) %>%
#   select(-id) %>%
#   merge(z2$nodes_df %>% rename(to = label) %>% select(id, to), by="to") %>%
#   mutate(to = id) %>%
#   arrange(to,from) -> edge_df_mod
# 
# 
# # z$nodes_df
# z2$nodes_df <- z2$nodes_df %>% mutate(label = recode_nodes(label))
# z2$nodes_df$fontname <- "Helvetica"
# z2$nodes_df$shape <- "ellipse"
# # z2$nodes_df$fillcolor <- ifelse(z2$nodes_df$id==13, "grey","white")
# z2$nodes_df$fixedsize <- FALSE
# z2$edges_df$penwidth <- abs(edge_df_mod$strength)*2
# z2$edges_df$penwidth <- 2
# # z$edges_df$label <- round(edge_df_mod$value,2)
# # z$edges_df$color <- ifelse(edge_df_mod$value > 0, "blue","red")
# z2$edges_df$color <- map_to_brewer(edge_df_mod$strength, bins=9, option="RdPu")
# # z2$edges_df$color <- map_to_viridis(edge_df_mod$strength, option="magma")
# 
# render_graph(z2)
```

```{r, echo=F, warning=F, message=F}
# boot_graph_legend <- (tibble(
#   Probability = seq(min(edge_df_mod$strength), max(edge_df_mod$strength), length=11),
# ) %>% 
#   ggplot(aes(y=1, x = Probability, fill = Probability)) +
#   geom_tile() +
#   scale_fill_distiller(palette="RdPu", direction=1)
# ) %>% 
#   cowplot::get_legend()
# 
# cowplot::ggdraw(boot_graph_legend)
```

Next, we can ask about the directions of the relationships within the model by examining the “completed partially directed acyclic graph” (CPDAG) derived from the structure learning algorithm. This is a graph with both directed and undirected edges which represents the “Markov equivalence class” of DAGs: the set of DAGs representing equivalent patterns of conditional (in)dependence among variables. Alternatively, this can be conceived of as “score equivalence,” the set of graphs which would produce equivalent scores for the observed data (Chickering, 2013). The CPDAG derived from our fit model contains 14 directed edges and 15 undirected arcs—indicating that the data do not constrain the directions of 15 of the edges in the final model. Of these undirected arcs, 10 directions are determined by our generative model constraints that we applied before fitting the model. Five arcs were determined neither by these constraints nor the data: these were $\text{holistic balance} \leftrightarrow \text{naturalism}$; $\text{medical skepticism} \leftrightarrow  \text{parental expertise}$, $\text{vaccine toxicity} \leftrightarrow \text{vaccine efficacy}$, $\text{vaccine toxicity} \leftrightarrow \text{vaccine danger}$ and $\text{vaccine danger} \leftrightarrow \text{vaccine efficacy}$. Thus, we feel that no strong conclusions should be drawn surrounding the counterintuitive relationships regarding vaccine toxicity, danger, and effectiveness.

@@

@@

```{r main_model_plot, echo=F, fig.width=8, fig.height=5, fig.align='center', fig.cap="Bayesian network model of intuitive theory surrounding vaccination decisions. Edge color indicates the magnitude of the generative (blue) or preventive (red) coeffient for the links."}

@@

# export graph image and render
export_graph(z, "../local/main-graph.pdf")
render_graph(z)
```

```{r, echo=F, warning=F, message=F}
## NOTE: I just can't figure out how to nicely display the graph and legend in code
## it's easier just to export them both and then combine with a graphics app

@@

cowplot::save_plot("../local/graph-legend.svg", main_graph_legend)
cowplot::ggdraw(main_graph_legend)

@@

@@

Finally, we can also compare the quality of our model predictions to the test-retest reliability of the belief scales, establishing an overall upper-bound on the predictability of these measures. As shown in table X, the Bayesian network model did approximately as well as possible for seven beliefs, but fell short for the other seven. These failures suggest that the model is missing the information that would be needed to predict these beliefs. Although the notion of an intuitive theory presupposes that there is some collective unit of cognitive machinery defining people’s thinking about a particular domain, defining the boundaries of an intuitive theory is far from straightforward. It could be that new measurements are needed to fully capture these aspects of the intuitive theory, or these beliefs are sufficiently peripheral that they should not be considered part of the intuitive theory. Given the fuzzy nature of the problem, and of cognition more generally, we expect that these types of decisions will be driven as much by investigators' pragmatic concerns as they are by psychological facts.

```{r, echo=FALSE, message=F, warning=F}
# compare S1 model predictions to test-retest reliability?

@@

kbl(model_fits_print, booktabs = T,
    digits = 3, 
    align = c('l','c','c','c'),
    # format = "latex",
    caption = "Correlations for observed and predicted out-of-sample values") %>% 
  kable_classic(full_width = F, html_font = "Cambria")
```
## Estimating the strength of evidence from changes in belief reports

Returning to the DAG shown in Figure 1 above, consider the subgraph $FIRE \rightarrow SMOKE$, capturing the fact that fire causes smoke. This relationship is represented as a probabilistic function, where the probability of SMOKE depends on the probability of FIRE. If we observe the presence of SMOKE, we can draw inferences about FIRE.

However, sometimes reasoners have only indirect evidence at their disposal. For instance, a person might not directly observe smoke, but instead hear from a witness who claimed to have seen it. Indirect or partial observations can be captured by augmenting the network, creating the graph $FIRE \rightarrow SMOKE \rightarrow WITNESS$. Here, a new node WITNESS represents another person’s testimony that they saw smoke, providing imperfect evidence for SMOKE, encoded by another probabilistic function. Observing WITNESS informs beliefs about SMOKE, in turn informing beliefs about FIRE. This approach to representing uncertain evidence is known as “virtual evidence” (Pearl, 1988).

To model the effects of participants observing evidence, we augmented the Bayesian network representing the cognitive model (shown in figure #) with an additional node representing the evidence. For instance, to model the effects of the “disease risk” intervention (Study 2), we added an evidence node as a child of “disease severity,” acting as virtual evidence for this node. To compute the CPT for this node as a function of disease severity beliefs, we estimated an _evidence ratio_ based on participants’ pretest and posttest credences in disease severity. To so do, we made use of the log-odds expression of Bayes rule,

@@

Where x was a binary variable representing the presence (1) or absence (0) of evidence at posttest, and participants’ prior and posterior (pretest and posttest) credences were transformed by the logit function. This allowed us to estimate the log evidence ratio implied by participants’ reactions to the intervention (ER1) as well as the intervening time period without any explicit evidence (ER0). Then, we constructed a CPT consistent with ER1. For instance, with an evidence ratio below 1, we set:

@@

To do this, we drew on a variety of sources, including a qualitative survey with participants in a previous study who had self-identified as vaccine skeptics. Out of 23 people invited to participate, n=16 participants completed the survey via Amazon Mechanical Turk (MTurk). All participants answered “yes” to the question “Are you concerned that childhood vaccinations might cause serious side-effects?” When asked, “Do you have concerns about all vaccines or only some specific vaccines?” nearly all participants (14 of 16) expressed being concerned about specific vaccines; 3 of these participants, as well as the remaining 2 participants, expressed concern about all vaccines.

@@

Drawing on these responses as well as anti-vaccine websites and academic articles on anti-vaccine skepticism (e.g., CITE), we generated a list of 13 underlying beliefs that might influence people’s decisions about whether to vaccinate their children (_vaccine intentions_, the last of our 14 beliefs). These included a variety of claims about vaccines, including beliefs about (1) the overall danger of vaccines (_vaccine danger_); (2) _toxic additives in vaccines_; and (3) _vaccine effectiveness_, how effective vaccines are in preventing disease; as well as a variety of specific claims about childhood diseases like measles, mumps, and rubella, including beliefs about (4) _disease rarity_ and (5) _disease severity_. Inspired by the many participants who invoked concerns about the effect of vaccines on infants’ immature immune systems, we also developed scales to assess beliefs that (6) the infant immune system is weak (_IIS: weakness_); (7) the infant immune system is limited in its capacity and can be easily overwhelmed (_IIS: limited capacity_); and (8) vaccines strain the infant immune system (_IIS: vaccines strain_). Beyond this, we developed two scales to assess general theories about parenting, including (9) general _parental protectiveness_; (10) _parental expertise_, namely the belief that parents usually know more about their children’s health than medical experts; and one scale to assess (12) _medical skepticism_, including concerns about pharmaceutical companies and corruption in the medical community. Finally, we decided to assess two broad worldviews: (12) _naturalism_, a general preference for natural over artificial things; and (13) _holistic balance_, one important aspect of attitudes toward alternative medicine (e.g., “The body is essentially self-healing and the task of a health care provider is to assist the healing process”; McFadden, Hernandez, & Ito, 2010).

_Holistic balance_ was the only belief for which we could identify a pre-existing psychometrically valid scale (McFadden et al., 2010). For each of the other 13 beliefs, we translated our qualitative insights into new psychometrically-valid scales through an iterative series of pilot studies not described here. We stipulated that each scale should be brief, composed of 4-6 statements for participants to evaluate; include at least one reverse-coded item; and be highly reliable (Cronbach’s α ≥ .80).

@@

_Method._ Participants in all of these studies were asked to respond at least once to 14 belief scales measuring beliefs relevant to vaccination decisions (as previously described). Across all studies, these scales were each presented on a separate survey page. On each page, participants were asked to rate “how much you agree or disagree” on a 7-point scale from Strongly disagree (coded as -3) to Strongly agree (+3). Scales, and questions within each scale, were presented in a random order for each participant. Two to four attention checks (“Please select somewhat agree” and “Please select somewhat disagree”) were embedded within two to four randomly chosen scales.

_Data processing._ For each scale, we generated a score for each participant by averaging their responses to each question in that scale (after reverse-coding where appropriate), and then rescaling these values to fall between 0 and 1, to represent a credence.

@@

Figure X presents correlations among the 14 belief scales measured in Study 1. As the figure shows, there are strong correlations among many of the beliefs, suggesting they are related by an underlying intuitive theory.

```{r, echo=FALSE, message=F, fig.height=7, fig.width=7, fig.align='center', fig.cap = "Correlations among 14 the belief scales measured in Study 1. Correlation matrix is arranged by hierarhical clustering algorithm to highlight connections among beliefs."}
# plot correlation matrix
library(corrplot)

@@

We conducted a mixed-effects Bayesian ordinal regression to assess the effects of the disease risk intervention on responses to the _Vaccination Intentions_ scale items, using the “brms” package for R (CITE). We regressed responses on phase (Phase 1 vs. 2, dummy-coded with Phase 1 as the baseline), condition (dummy-coded with the Disease Risk condition as the baseline, so as to assess difference scores in our condition of primary interest), and the interaction between phase and condition. We included random intercepts by subject and scale item. In this model, the unique effect of the intervention on vaccination intention responses is assessed by the interaction term (indicating the unique effect of Phase 2 for intervention versus no-intervention participants). The disease risk intervention increased participants’ vaccination intention ratings (see Table X). We consider this to be a full replication of Horne, Powell, et al.’s (XX) results.

@@

Figure X (right) shows the correlation between predicted and observed belief changes conditional on the disease risk intervention.

```{r, echo=FALSE}
# create cognitive model for predicting change scores
base_model <- create_bnlearn_cog_model(main_model, train)
```

@@

@@

1) Where beliefs about two states of affairs, A and B, are related through a shared intuitive theory (e.g., A causes B), those beliefs will be systematically correlated across individuals (e.g., people who believe A will also tend to believe B);

>>>>

---
title: "Supplemental Materials"
subtitle: "Untitled vaccination intuitive theory paper"
author: "Derek Powell, Kara Weisman, and Ellen M. Markman"
date: "Jul 17, 2021 (DRAFT)"
output:
  pdf_document: 
    latex_engine: xelatex
    fig_caption: yes
bibliography: references.bib
csl: pnas.csl
figsintext : yes
always_allow_html: true
header-includes:
 \usepackage{float}
---

@@

Figure \@ref(fig:dag-example) shows an example of a simple Bayesian Network. The network itself is a _directed acyclic graph_ (DAG), a type of graphical model. Graphical models represent variables as nodes connected by edges. In a DAG, these edges are directed, flowing from parent to child, and are constrained so that they cannot form cycles or loops. In a Bayesian network, these directed edges encode conditional dependencies among variables, so that the conditional probability distribution for a given variable is defined only in terms of its direct parents. Together, a child node and its parents are called a _family_.

```{r dag-example, echo=F, fig.align='center', out.width = '40%', fig.cap="Example DAG illustrating some variables related to wildfires."}

@@

In this graph, the edge $FIRE \rightarrow SMOKE$ captures that fire causes smoke.^[The example of cause and effect is especially intuitive, but these edges need not be causal. Instead, they are better thought of as representing “influence relations”, which may also include relationships like logical implication and set membership (Williamson, 2001).] This relationship is represented as a probabilistic function, where the probability of smoke depends on the probability of fire. The probability of a node Y can be defined from its parents X in many different ways. For binary nodes, the most flexible specification is a conditional probability table (CPT) with 2^n^ entries specifying the probability of Y given the entire joint probability distribution over X, where n is the number of parents X. Given the exponential complexity of the CPT, simpler functions are often used to specify the probability of the child node, such as the noisy-or and noisy-and-not functions. A noisy-or function assumes each of the parents are independent generative forces, each with a probability of generating the effect. A noisy-and-not function assumes each of the parents are independent preventive forces, specifying the probability of the child node as the probability that it is generated _and not_ prevented by any one of its possible preventers. We construct our cognitive model parameterization using a combination of these two functions, using a distribution sometimes referred to as a “DeMorgan gate” [@maaskant.druzdzel2008]. In our model, the probability of the child node Y can be expressed as:

@@

A cognitive model composed of these types of relations is capable of capturing a causal system of non-interactive generative and preventive causes. Although people’s intuitive theories are likely more flexible than this, there is ample evidence that people are highly capable of reasoning according to these types of causal relationships [see @holyoak.cheng2011 for a review], and moreover, that their causal learning is strongly biased toward learning these kinds of relations [e.g. @novick.cheng2004].

@@

To conduct an efficient search through the super-exponential space of DAGs, we tested several variants of two structure learning algorithms implemented in the “bnlearn” R package [@scutari2010a]. The first is a purely score-based “hill-climbing” algorithm  and the second was a hybrid algorithm that combined the hill-climbing algorithm with a prior “restriction” phase that pruned down the space of DAGs. This pruning was accomplished with the “min-max parents and children” algorithm, which can be tuned with a hyperparameter alpha, which we tested at several levels. To speed computations, we restricted the maximum number of parents in the graph to five for all nodes, as the complexity of the scoring function is exponential in the number of parents. In all resulting networks no node was found to have more than four parents, suggesting this did not impact the final results.

@@

Figure \@ref(fig:cross-validation) presents the results of these cross validation tests. As seen in the figure, all structure learning approaches produced models with similar out-of-sample performance, with more variability among folds and runs than between algorithms. Thus the cross-validation does not uniquely identify an optimal structure learning strategy.

@@

```{r cross-validation, echo=FALSE, fig.align='center', fig.height=3, fig.width=4, fig.cap="Cross-validation performance for diffrent structure learning algorithms on the training data from Study 1."}

@@

We used a bootstrapping procedure to estimate the confidence in each of the links in this model, conducting the structure learning procedure repeatedly for 200 bootstrapped resamples of the original training data. Figure \@ref(fig:disp-bootgraph) visualizes the frequency with which each arc appeared in the resulting model across these replicates, and figure \@ref(fig:disp-maingraph) visualizes the model with the coefficients for each link encoded by color.

@@

@@

```{r boot_strength_plot, include=F, fig.width=5, fig.height=8}
# 
hyde_str <- df_to_hyde_string(boot_res %>% filter(strength>.2, direction>0) %>% select(from,to))
h2 <- HydeNetwork(as.formula(hyde_str))
# plot(h2)

z2 <- custom_plot.HydeNetwork(h2)

boot_res %>%
  filter(direction>0, strength>.2) %>%
  # filter(parameter!="phi",parameter!="w_leak") %>%
  # rename(from = parameter) %>%
  # mutate(from = gsub("w_","",from)) %>%
  mutate(strength = strength*direction) %>%
  select(to, from, strength) %>%
  merge(z2$nodes_df %>% rename(from = label) %>% select(id, from), by="from") %>%
  mutate(from = id) %>%
  select(-id) %>%
  merge(z2$nodes_df %>% rename(to = label) %>% select(id, to), by="to") %>%
  mutate(to = id) %>%
  arrange(to,from) -> edge_df_mod

# z$nodes_df
z2$nodes_df <- z2$nodes_df %>% mutate(label = recode_nodes(label))
z2$nodes_df$fontname <- "Helvetica"
z2$nodes_df$shape <- "ellipse"
z2$nodes_df$fillcolor <- ifelse(z2$nodes_df$label=="Vaccination\nIntentions", "grey","white")
z2$nodes_df$fixedsize <- FALSE
z2$edges_df$penwidth <- abs(edge_df_mod$strength)*2
z2$edges_df$penwidth <- 1.5
# z$edges_df$label <- round(edge_df_mod$value,2)
# z$edges_df$color <- ifelse(edge_df_mod$value > 0, "blue","red")
z2$edges_df$color <- map_to_brewer(edge_df_mod$strength, bins=9, option="RdPu")
# z2$edges_df$color <- map_to_viridis(edge_df_mod$strength, option="magma")
# 
# render_graph(z2)

export_graph(z2, "../local/boot-graph.pdf")
```

```{r, include=F}
boot_graph_legend <- (tibble(
  Probability = seq(min(edge_df_mod$strength), max(edge_df_mod$strength), length=11),
) %>%
  ggplot(aes(y=1, x = Probability, fill = Probability)) +
  geom_tile() +
  scale_fill_distiller(palette="RdPu", direction=1)
) %>%
  cowplot::get_legend()

cowplot::save_plot("../local/boot-graph-legend.svg", boot_graph_legend)
# cowplot::ggdraw(boot_graph_legend)
```

```{r disp-bootgraph, echo=F, fig.width=8, fig.height=5, fig.cap="Bootstrapped confidence estimates for Bayesian network edges. Edge color indicates frequency with which edges appeared when conducting search across 200 bootstrap replicates. Only edges appearing in at least 20\\% of graphs are shown."}

## assemble bootstrapping graph and legend into pdf and display here
knitr::include_graphics("../local/boot-graph.pdf") 
```

A Bayesian network is a factorized representation of a probability distribution and any given network is a member of a "Markov equivalence class"—a set of networks capable of factorizing the same distribution. This set of DAGs can be represented as a "completed partially directed acyclic graph" (CPDAG), a graph with both directed and undirected edges. The equivalence class for our model can be represented as a CPDAG with 14 directed edges and 15 undirected edges—indicating that the data do not constrain the directions of 15 of the edges in the final model. However, 10 of these edges’ directions are directly constrained by our top-down "generative model" constraints, which in turn constrains all but two of the remaining undirected edges, including the counterintuitive edges running toward vaccine toxins. This is also reflected in a bootstrapping analysis applying our chosen structure learning algorithm against 1000 bootstrapped replicates of our data. We found that the direction of some edges within the graph differed across bootstrapped runs (Holistic balance -- naturalism, medical skepticism -- parental expertise), but that these unintuitive relationships surrounding vaccine toxins were robust.

Altogether, these findings might be seen as demonstrating the need for a heavier hand in establishing top-down constraints on the model fitting. Certainly, they call for caution in interpreting the relationships represented in this model. Due to their tight connections with structural causal networks, it is tempting to interpret the edges in Bayesian networks as “causal.” However, these relationships are best thought of as probabilistic dependencies or “influence” relations (Williamson, 2001). Indeed, there are other cases where it is unclear if an edge should be directed at all: e.g., it is not obvious to us what would be the “correct” direction for the edge between naturalism and holistic balance. We might even speculate there is an additional unmeasured variable that connects these beliefs and that would best account for their association. Ultimately, this model should be understood as offering a plausible, but not definitive, account for a cognitive model connecting these different beliefs.

@@

The three nodes with direct connections to _vaccine intentions_ are beliefs about _vaccine effectiveness_, _vaccine danger_, and the severity of childhood diseases (_disease severity_)---concrete beliefs that are, at face value, especially closely related to vaccination decisions. The directionality of these edges is informed by the constraints we applied via the generative model principle, but the identification of these three factors as the most proximal to vaccine intentions is a product of the structure learning approach. This gives us some confidence that the model is capturing important relationships.

Other parts of the model shed new light on the role of lay theories in vaccine decisions. For example, _naturalism_—--the general view that natural things are better than artificial things—appears to be strongly related to _medical skepticism_ and _parental expertise_; all three of these abstract beliefs are related to concrete beliefs that, in turn, feed into participants’ vaccination intentions. This hierarchical structure is at least partly a product of our constraints that sorted these beliefs into “tiers” of abstractness. However, the fact that a model with these constraints still functions well for prediction demonstrates that this hierarchical structure is plausible.

Finally, this model structure also helps explain some initially surprising findings. For example, before collecting these data we speculated that dispelling the misconceptions about the capacity of the infant immune system could promote positive attitudes toward vaccination. We were disappointed to observe the weak first-order correlation between this belief and vaccine intentions in our behavioral data ($r$ = -.09 in our training split). The model sheds light on this surprising (lack of) relationship: Although the belief that the infant immune system is limited in capacity is positively related to the belief that vaccines strain the immune system—discouraging vaccination, as we had assumed—it also seems to promote the belief that childhood diseases have severe consequences for young children (_disease severity_), which might, in turn, encourage vaccination. These countervailing forces—which, on average, seem to have canceled each other out in this sample—dissuaded us from our plans to develop an intervention aimed at dispelling misconceptions about limited capacity in order to encourage childhood vaccination.

@@

```{r main_model_plot, include=F}

@@

# export graph image
export_graph(z, "../local/main-graph.pdf")
# render_graph(z)
```

```{r, include=F}
## NOTE: I just can't figure out how to nicely display the graph and legend in code
## it's easier just to export them both and then combine with a graphics app

@@

cowplot::save_plot("../local/graph-legend.svg", main_graph_legend)
# cowplot::ggdraw(main_graph_legend)

@@

```{r disp-maingraph, echo=F, fig.width=8, fig.height=5, fig.align='center', fig.cap="Bayesian network model of intuitive theory surrounding vaccination decisions. Edge color indicates the magnitude of the generative (blue) or preventive (red) coeffient for the links."}

## assemble main model graph and legend into pdf and display here (main-graph-combined.pdf)
## currently just the output graph w/ no legend (main-graph.pdf)
knitr::include_graphics("../local/main-graph.pdf") 
```

@@

Finally, we can also compare the quality of our model predictions to the test-retest reliability of the belief scales (estimated from the pre-test and post-test measures in the control condition of Study 2), establishing an overall upper-bound on the predictability of these measures. As shown in Table \@ref(tab:model_valid_table), the Bayesian network model did approximately as well as possible for seven beliefs, but fell short for the other seven. These failures suggest that the model is missing the information that would be needed to predict these beliefs. Although the notion of an intuitive theory presupposes that there is some collective unit of cognitive machinery defining people’s thinking about a particular domain, defining the boundaries of an intuitive theory is far from straightforward. It could be that new measurements are needed to fully capture these aspects of the intuitive theory, or these beliefs are sufficiently peripheral that they should not be considered part of the intuitive theory. Given the fuzzy nature of the problem, and of cognition more generally, we expect that these types of decisions will be driven as much by investigators' pragmatic concerns as they are by psychological facts.

```{r model_valid_table, echo=FALSE, message=F, warning=F}
# compare S1 model predictions to test-retest reliability?

@@

kbl(model_fits_print, booktabs = T,
    digits = 3, 
    align = c('l','c','c','c'),
    # format = "latex",
    caption = "Correlations for observed and predicted out-of-sample values") %>% 
  kable_classic(full_width = F, html_font = "Cambria")
```

## Estimating the strength of evidence from changes in belief reports

Returning to the DAG shown in Figure \@ref(fig:dag_example) above, consider the subgraph $FIRE \rightarrow SMOKE$, capturing the fact that fire causes smoke. This relationship is represented as a probabilistic function, where the probability of SMOKE depends on the probability of FIRE. If we observe the presence of SMOKE, we can draw inferences about FIRE.

However, sometimes reasoners have only indirect evidence at their disposal. For instance, a person might not directly observe smoke, but instead hear from a witness who claimed to have seen it. Indirect or partial observations can be captured by augmenting the network, creating the graph $FIRE \rightarrow SMOKE \rightarrow WITNESS$. Here, a new node WITNESS represents another person’s testimony that they saw smoke, providing imperfect evidence for SMOKE, encoded by another probabilistic function. Observing WITNESS informs beliefs about SMOKE, in turn informing beliefs about FIRE. This approach to representing uncertain evidence is known as “virtual evidence” [@pearl1988].

To model the effects of participants observing evidence, we augmented the Bayesian network representing the cognitive model (shown in Figure \@ref(fig:disp-maingraph)) with an additional node representing the evidence. For instance, to model the effects of the “disease risk” intervention (Study 2), we added an evidence node as a child of “disease severity,” acting as virtual evidence for this node. To compute the CPT for this node as a function of disease severity beliefs, we estimated an _evidence ratio_ based on participants’ pretest and posttest credences in disease severity. To so do, we made use of the log-odds expression of Bayes rule,

@@

Where $x$ was a binary variable representing the presence (1) or absence (0) of evidence at posttest, and participants’ prior and posterior (pretest and posttest) credences were transformed by the logit function. This allowed us to estimate the log evidence ratio implied by participants’ reactions to the intervention ($ER_1$) as well as the intervening time period without any explicit evidence ($ER_0$). Then, we constructed a CPT consistent with $ER_1$. For instance, with an evidence ratio below 1, we set:

@@

To do this, we drew on a variety of sources, including a qualitative survey with a set of pre-screened participants who had self-identified as vaccine skeptics. Out of 23 people invited to participate, n=16 participants completed the survey via Amazon Mechanical Turk (MTurk). All participants had previously answered “yes” to the question “Are you concerned that childhood vaccinations might cause serious side-effects?” When asked, “Do you have concerns about all vaccines or only some specific vaccines?” nearly all participants (14 of 16) expressed being concerned about specific vaccines; 3 of these participants, as well as the remaining 2 participants, expressed concern about all vaccines.

@@

Drawing on these responses as well as anti-vaccine websites and academic articles on anti-vaccine skepticism [@dube.etal2015, @larson.etal2015, @salmon.etal2015, @williams2014], we generated a list of 13 underlying beliefs that might influence people’s decisions about whether to vaccinate their children (_vaccine intentions_, the last of our 14 beliefs). These included a variety of claims about vaccines, including beliefs about (1) the overall danger of vaccines (_vaccine danger_); (2) _toxic additives in vaccines_; and (3) _vaccine effectiveness_, how effective vaccines are in preventing disease; as well as a variety of specific claims about childhood diseases like measles, mumps, and rubella, including beliefs about (4) _disease rarity _and (5) _disease severity_. Inspired by the many participants who invoked concerns about the effect of vaccines on infants’ immature immune systems, we also developed scales to assess beliefs that (6) the infant immune system is weak (_IIS: weakness_); (7) the infant immune system is limited in its capacity and can be easily overwhelmed (_IIS: limited capacity_); and (8) vaccines strain the infant immune system (_IIS: vaccines strain_). Beyond this, we developed two scales to assess general theories about parenting, including (9) general _parental protectiveness_; (10) _parental expertise_, namely the belief that parents usually know more about their children’s health than medical experts; and one scale to assess (12) _medical skepticism_, including concerns about pharmaceutical companies and corruption in the medical community. Finally, we decided to assess two broad worldviews: (12) _naturalism_, a general preference for natural over artificial things; and (13) _holistic balance_, one important aspect of attitudes toward alternative medicine (e.g., “The body is essentially self-healing and the task of a health care provider is to assist the healing process”; [@mcfadden.etal2010].

_Holistic balance_ was the only belief for which we could identify a pre-existing psychometrically valid scale [@mcfadden.etal2010]. For each of the other 13 beliefs, we translated our qualitative insights into new psychometrically-valid scales through an iterative series of pilot studies not described here. We stipulated that each scale should be brief, composed of 4-6 statements for participants to evaluate; include at least one reverse-coded item; and be highly reliable (Cronbach’s $\alpha \geq .80$).

@@

### Method

Participants in all of these studies were asked to respond at least once to 14 belief scales measuring beliefs relevant to vaccination decisions (as previously described). Across all studies, these scales were each presented on a separate survey page. On each page, participants were asked to rate “how much you agree or disagree” on a 7-point scale from Strongly disagree (coded as -3) to Strongly agree (+3). Scales, and questions within each scale, were presented in a random order for each participant. Two to four attention checks (“Please select somewhat agree” and “Please select somewhat disagree”) were embedded within two to four randomly chosen scales.

### Data processing

For each scale, we generated a score for each participant by averaging their responses to each question in that scale (after reverse-coding where appropriate), and then rescaling these values to fall between 0 and 1, to represent a credence.

@@

Figure \@ref(fig:corrplot) presents correlations among the 14 belief scales measured in Study 1. As the figure shows, there are strong correlations among many of the beliefs, suggesting they are related by an underlying intuitive theory.

```{r corrplot, echo=FALSE, message=F, fig.height=7, fig.width=7, fig.align='center', fig.cap = "Correlations among 14 the belief scales measured in Study 1. Correlation matrix is arranged by hierarhical clustering algorithm to highlight connections among beliefs."}
# plot correlation matrix
library(corrplot)

@@

We conducted a mixed-effects Bayesian ordinal regression to assess the effects of the disease risk intervention on responses to the _Vaccination Intentions_ scale items, using the “brms” package for R [@burkner2017]. We regressed responses on phase (Phase 1 vs. 2, dummy-coded with Phase 1 as the baseline), condition (dummy-coded with the Disease Risk condition as the baseline, so as to assess difference scores in our condition of primary interest), and the interaction between phase and condition. We included random intercepts by subject and scale item. In this model, the unique effect of the intervention on vaccination intention responses is assessed by the interaction term (indicating the unique effect of Phase 2 for intervention versus no-intervention participants). The disease risk intervention increased participants’ vaccination intention ratings (see Table \@ref(tab:s2_model)). We consider this to be a full replication of Horne et al.’s [-@horne.etal2015] results.

@@

Figure \@ref(fig:s2_plot) (right) shows the correlation between predicted and observed belief changes conditional on the disease risk intervention.

```{r s2_plot, echo=FALSE}
# create cognitive model for predicting change scores
base_model <- create_bnlearn_cog_model(main_model, train)
```

@@

The models’ predictions are strongly correlated with observed changes. Thus the model developed in Study 1 generalizes both across samples and across tasks: The model was developed from the correlations among beliefs in one sample of participants at a single time-point, and was able to accurately predict how a separate sample of participants revised their beliefs following evidence.

@@

1. Where beliefs about two states of affairs, A and B, are related through a shared intuitive theory (e.g., A causes B), those beliefs will be systematically correlated across individuals (e.g., people who believe A will also tend to believe B);

2. When two beliefs are related by an intuitive theory, evidence affecting one of those beliefs will also affect the other in accordance with their relationship in the intuitive theory (e.g., evidence for B will increase credence in A, and vice versa); and
